---
title: 'Homework 4: Bags, Forests, Boosts, oh my'
author: "Skylar White"
date: "2/28/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(ISLR)
library(tree)
library(caret)
library(stringr)
library(rpart)
library(rpart.plot)
library(partykit)
theme_set(theme_bw())
theme_set(theme_bw())
```

## Problem 1

Problem 7 from Chapter 8 in the text. To be specific, please use a sequence of
`ntree` from 25 to 500 in steps of 25 and `mtry` from 3 to 9 for by 1. 

In the lab, we applied random forests to the Boston data using mtry=6
and using ntree=25 and ntree=500. Create a plot displaying the test
error resulting from random forests on this data set for a more comprehensive range of values for mtry and ntree. You can model your
plot after Figure 8.10. Describe the results obtained.
## Answer 1
```{r}
library(MASS)
library(randomForest)
set.seed(1)
train = sample(1:nrow(Boston), nrow(Boston)/2)
boston.test=Boston[-train,"medv"]
```

```{r}
set.seed(1)
rf.boston= randomForest(medv~.,data=Boston, subset=train,
mtry=6, importance =TRUE)
yhat.rf = predict(rf.boston ,newdata=Boston[-train,])
mean((yhat.rf-boston.test)^2)
```
```{r}
rf.boston
```
```{r}
set.seed(1234)
df <- tbl_df(Boston)
inTraining <- createDataPartition(df$medv, p = .75, list = F)
training <- df[inTraining, ]
testing  <- df[-inTraining, ]

results <- data_frame(ntree = seq(25, 500, by = 25),
                      mtry_3 = rep(NA, length(ntree)),
                      mtry_4 = rep(NA, length(ntree)),
                      mtry_5 = rep(NA, length(ntree)),
                      mtry_6 = rep(NA, length(ntree)),
                      mtry_7 = rep(NA, length(ntree)),
                      mtry_8 = rep(NA, length(ntree)),
                      mtry_9 = rep(NA, length(ntree)))

for (i in 1:20){
rf_boston_cv <- train(medv~., 
                      data = training,
                      method = "rf",
                      ntree = results$ntree[i],
                      importance = T,
                      tuneGrid = data.frame(mtry = 3:9))
results[i, 2:8] <- rf_boston_cv$results$RMSE^2}

```

```{r}
ggplot(data = results, 
            aes(x = ntree)) + 
  geom_line(aes(y = mtry_3), color = "black")+
  geom_line(aes(y = mtry_4), color = "blue") +
  geom_line(aes(y = mtry_5), color = "orange")+
  geom_line(aes(y = mtry_6), color = "yellow")+
  geom_line(aes(y = mtry_7), color = "green")+
  geom_line(aes(y = mtry_8), color = "purple")+
  geom_line(aes(y = mtry_9), color = "red")+
  ylab(label="Test Error Rate") + 
  xlab("Number of Trees")
```

## Problem 2

Problem 8 from Chapter 8 in the text. Set your seed with 9823 and split into 
train/test using 50\% of your data in each split. In addition to 
parts (a) - (e), do the following:

1. Fit a gradient-boosted tree to the training data and report the estimated 
test MSE. 
2. Fit a multiple regression model to the training data and report the 
estimated test MSE
3. Summarize your results. 

In the lab, a classification tree was applied to the Carseats data set after converting Sales into a qualitative response variable. Now we will
seek to predict Sales using regression trees and related approaches,
treating the response as a quantitative variable.
(a) Split the data set into a training set and a test set.
```{r}
set.seed(9823)
df <- tbl_df(Carseats)
train <- sample(1:nrow(Carseats), nrow(Carseats) / 2)
training <- Carseats[train, ]
testing <- Carseats[-train, ]
```

(b) Fit a regression tree to the training set. Plot the tree, and interpret the results. What test error rate do you obtain?
```{r}
carseats_tree <- tree(Sales ~ ., data = training)
summary(carseats_tree)
```
```{r}
plot(carseats_tree)
text(carseats_tree, pretty = 0)
```

```{r}
yhat <- predict(carseats_tree, newdata = testing)
mean((yhat - testing$Sales)^2)
```

(c) Use cross-validation in order to determine the optimal level of
tree complexity. Does pruning the tree improve the test error
rate?
```{r}
cv_carseats <- cv.tree(carseats_tree)
plot(cv_carseats$size, cv_carseats$dev, type = "b")
tree.min <- which.min(cv_carseats$dev)
```
```{r}
prune_carseats <- prune.tree(carseats_tree, best = 8)
plot(prune_carseats)
text(prune_carseats, pretty = 0)
```
```{r}
yhat <- predict(prune_carseats, newdata = testing)
mean((yhat - testing$Sales)^2)
```
No, it increases it.

(d) Use the bagging approach in order to analyze this data. What
test error rate do you obtain? Use the importance() function to
determine which variables are most important.
334 8. Tree-Based Methods

```{r}
bag_carseats <- randomForest(Sales ~ ., data = training, mtry = 10, ntree = 500, importance = TRUE)
yhat_bag <- predict(bag_carseats, newdata = testing)
mean((yhat_bag - testing$Sales)^2)
```
```{r}
importance(bag_carseats)
```

(e) Use random forests to analyze this data. What test error rate do
you obtain? Use the importance() function to determine which
variables are most important. Describe the effect of m, the number of variables considered at each split, on the error rate
obtained.
```{r}
rf_carseats <- randomForest(Sales ~ ., data = training, mtry = 3, ntree = 500, importance = TRUE)
yhat_rf <- predict(rf_carseats, newdata = testing)
mean((yhat_rf - testing$Sales)^2)
```
Having m= root(p) makes the error rate relatively high compared with bagging.

```{r}
importance(rf_carseats)
```
1. Fit a gradient-boosted tree to the training data and report the estimated 
test MSE. 
```{r}
set.seed(9823)
grid <- expand.grid(interaction.depth = c(1, 3), 
                    n.trees = seq(0, 2000, by = 100),
                    shrinkage = c(.01, 0.001),
                    n.minobsinnode = 10)
trainControl <- trainControl(method = "cv", number = 5)
gbm_Carseats <- train(Sales~., 
                    data = training, 
                    distribution = "gaussian", 
                    method = "gbm",
                    trControl = trainControl, 
                    tuneGrid = grid,
                    verbose = FALSE)
gbm_Carseats
```
```{r}
test_preds <- predict(gbm_Carseats, newdata = testing)
carseats_test_df <- testing %>%
  mutate(y_hat_gbm = test_preds,
         sq_err_gbm = (y_hat_gbm - Sales)^2)
mean(carseats_test_df$sq_err_gbm)
```

2. Fit a multiple regression model to the training data and report the 
estimated test MSE
```{r}
lm_carseats <- lm(Sales ~ ., data = training)
```
```{r}
test_preds <- predict(lm_carseats, newdata = testing)
carseats_test_df <- testing %>%
  mutate(y_hat_gbm = test_preds,
         sq_err_gbm = (y_hat_gbm - Sales)^2)
mean(carseats_test_df$sq_err_gbm)
```

3. Summarize your results. 
```{r}
summary(gbm_Carseats)
```
```{r}
summary(lm_carseats)
```

